---
title: "ES4"
author: "Martin Kosík"
date: '2022-06-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("plm")
# install.packages("texreg")
# install.packages("modelsummary")

library(plm)
library(texreg)
# library(modelsummary)
```

For more info on dynamic panel data models see Cameron and Trivedi, page 763 (22.5) or [Hansen](https://www.ssc.wisc.edu/~bhansen/econometrics/Econometrics.pdf) page 639 (663 in the pdf document).

## 1)

Load data and inspect them
```{r}
data("EmplUK", package = "plm")
head(EmplUK)
```

Run Arellano-Bond estimator
```{r}
a1 <- pgmm(log(emp) ~ lag(log(emp), 1:2) + lag(log(wage), 0:1)
          + lag(log(capital), 0:2)+lag(log(output), 0:2) | lag(log(emp), 2:9),
          data = EmplUK, effect = "twoways", model = "onestep")



a2 <- pgmm(log(emp) ~ lag(log(emp), 1:2) + lag(log(wage), 0:1)
           + lag(log(capital), 0:2)+lag(log(output), 0:2) | lag(log(emp), 2:9),
           data = EmplUK, effect = "twoways", model = "twosteps")


b <- pgmm(log(emp) ~ lag(log(emp), 1:2) + lag(log(wage), 0:1)
           + log(capital)+lag(log(output), 0:1) | lag(log(emp), 2:9),
           data = EmplUK, effect = "twoways", model = "twosteps")




screenreg(list(a1 = a1, a2 = a2, b = b),
          digits = 3, omit.coef = "(Intercept)")

#modelsummary(list(a1 = a1, a2 = a2, b = b),
#          digits = 3, omit.coef = "(Intercept)")


```

## 2)
This corresponds to Anderson-Hsiao estimator which is less efficient. We need at least 3 time periods. 
```{r}
AH <- pgmm(log(emp) ~ lag(log(emp), 1)+ lag(log(wage), 0:1) +
             lag(log(capital), 0:1) | lag(log(emp), 2),        
           data = EmplUK, effect = "twoways", model = "onestep")

summary(AH)

```
## 3)

In case of strictly exogenous regressors we can use any number of lag and leads of $x_{it}$ as instruments for $\Delta y_{it-1}$ whereas in case of pre-determined regressors we can use only the lags of $x_{it}$.
```{r}
# Arellano-Bond one-step estimator with predetermined regressors
GMM_DIF <- pgmm(log(emp) ~ lag(log(emp), 1)+ lag(log(wage), 0:1) +
             lag(log(capital), 0:1) | lag(log(emp), 2:9) +
             lag(log(wage), 2:9) + lag(log(capital), 2:9),        
           data = EmplUK, effect = "twoways", model = "onestep")
summary(GMM_DIF)


# Blundell-Bond one-step systems GMM estimator
GMM_SYS <- pgmm(log(emp) ~ lag(log(emp), 1)+ lag(log(wage), 0:1) +
                        lag(log(capital), 0:1) | lag(log(emp), 2:9) +
                        lag(log(wage), 2:9) + lag(log(capital), 2:9),        
                      data = EmplUK, effect = "twoways", model = "onestep", 
                      transformation = "ld")

summary(GMM_SYS, robust = TRUE)

```

