---
title: "Exercice session 1"
author: "Martin Kosík"
date: "May 23, 2022"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Fundamentals

### Packages

You can install new packages using `install.packages()` function. The installed packages are loaded with `library()`.

```{r warning=FALSE}
#install.packages("here")
library(here)
```

### Objects

Objects can be assigned with either `<-` or `=` operators. They have exactly the same effect, it is just the matter of style preference.

```{r}
string_object_1 <- "hello_world!"
string_object_2 = "hello_world!"

string_object_1
string_object_2
```

Objects can be removed with `rm()` function.

```{r}
rm(string_object_1)
#string_object_1
```

```{r}
rm(list = c('string_object_1', 'string_object_2'))
```

To remove all objects, use `rm(list=ls())`.

### Vectors

A sequence of objects of the same type (e.g., string, interger, float, boolean, factor) forms a vector.

```{r}
string_vector <- c("one", "two", "three")
integer_vector <- 1:3  # same as c(1, 2, 3)
float_vector <- c(1.0, 2.0, 3.0)
boolean_vector <- c(FALSE, TRUE, FALSE) # or alternatively c(F, T, F)
```

The n-th element of vector x is accessed by `x[n]`

```{r}
string_vector[2]
string_vector[c(2, 3)]

```

### Dataframes

The vectors of the same length can be combined into a Dataframe

```{r}
example_df <- data.frame(first_column = string_vector, second_column = integer_vector, third_column = float_vector)
example_df
```

There are several ways how to access a specific vector (column) of a dataframe

```{r}
example_df[ ,2]
example_df[ ,'second_column']
example_df$second_column

example_df$second_column

```

You can also access multiple columns passing a vector with column names, column indices, or boolean vector. The output is then a dataframe.

```{r}
example_df[3, 'third_column'] <- 5
example_df[3, 'third_column'] = 5

example_df[c(2,3), 'third_column'] <- c(4, 5)

example_df[ ,c(2, 3)]
example_df[ ,c('third_column', 'second_column')]
example_df[ ,c(FALSE, TRUE, TRUE)]

```

The same applies to accessing rows

```{r}
example_df[c(2, 3), ]
example_df[c(FALSE, TRUE, TRUE), ]

```

```{r}
example_df[example_df$first_column == 'two',]
```

### Lists

Lists are objects that can contain any kind of objects in them (strings, floats, booleans, another lists, ...)

```{r}
cool_list <- list(55, 'cool item', list(TRUE, 6.626262))
cool_list
```

You can access the elements by their index

```{r}
cool_list[1]

```

```{r}
cool_list[[1]]

```

Note the subtle difference between the two: in the first case the output is a list with a single element 55, whereas in the second case the output is only 55.

You can also name the elements of list and them access them by their name

```{r}
cool_named_list <- list('first el.' = 55, 
                        'second el.' = 'cool item',
                        'third el.' = list(TRUE, 6.626262))

```

```{r}
cool_named_list[['second el.']]

```

You can combine the two ways

```{r}
cool_named_list[['third el.']] [[2]]
```

## Basic econometrics in R

To estimate basic linear regression of form $y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \epsilon$ use `lm(y ~ x_1 + x_2, data)`.

```{r}
swiss
```

```{r}
data("swiss")


basic_regression <- lm(Infant.Mortality ~ Fertility + Education +  Catholic + Agriculture, data = swiss)

summary(basic_regression)
```

By default, `lm()` outputs standard errors assuming homoskedasticity. You almost never want to use them.

There are several ways how to obtain heteroskedasticity-robust SE.

First, `estimatr` package has `lm_robust()` function whose `se_type` argument can be either "classical", "HC0", "HC1" (stata), "HC2", or "HC3".

See [this](https://declaredesign.org/r/estimatr/articles/mathematical-notes.html) for the mathematical details.

```{r}
# install.packages('estimatr', repos = "http://cran.us.r-project.org")
library(estimatr)
fancy_regression <- lm_robust(Infant.Mortality ~ Fertility + Education +  Catholic + Agriculture, data = swiss, se_type = "HC1")
summary(fancy_regression)
```

Alternatively, in `fixest` package you can use `feols()` function and passing in `vcov = "hetero"` to get HC0 (White) SE. Additionally, you can also apply the small sample correction by passing in `ssc = ssc(adj = TRUE)`. Learn more [here](https://lrberge.github.io/fixest/articles/standard_errors.html)

```{r}
# install.packages('fixest', repos = "http://cran.us.r-project.org")
library(fixest)
even_fancier_regression <- feols(Infant.Mortality ~ Fertility + Education +  Catholic + Agriculture, data = swiss, vcov = "hetero", ssc = ssc(adj = TRUE))
summary(even_fancier_regression)

```

Both packages also support SE clustering (more on that in some future ES).

## Regression tables

For regression tables, I would recommend [modelsummary package](https://vincentarelbundock.github.io/modelsummary/articles/modelsummary.html) (it can output html, latex, markdown, or gt formats) but there are a lot of options out there. Stargazer package is also popular for output to latex but it is not maintained and there might not be compatibility with some model formats.

```{r eval=FALSE}
library(stargazer)
stargazer(basic_regression, basic_regression)
```

## Introduction to Tidyverse

The tidyverse is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures.

A common feature of most tidyverse is that they take the dataframe as the first argument.

Here is a comparison of a few common operations done in tidyverse and base R (taken from [here](https://dplyr.tidyverse.org/articles/base.html)). Also notice that

```{r warning=FALSE, message=FALSE}
library(tidyverse)
# Subseting dataframe
# Tidyverse
swiss_filter <- filter(swiss, Catholic < 5, Education > 10)
# Base R
swiss_filter <-	swiss[which(swiss$Catholic < 5), , drop = FALSE]


# Order by a column value
# Tidyverse
swiss_order <- arrange(swiss, Catholic)
# Base R
swiss_order <-	swiss[order(swiss$Catholic), , drop = FALSE]

# Create a new column
swiss_new_col <- mutate(swiss, new_col = Examination + Education)
# Base R
swiss_new_col <- swiss
swiss_new_col$new_col <- swiss_new_col$Examination + swiss_new_col$Education

```

Another important feature of Tidyverse is the use of pipes ( %\>% ). It does have only stylistic function (i.e., it makes the code more readable).

The pipe ( %\>% ) makes the object before it the first argument of the function after it. For example

```{r}
swiss_filter <- swiss %>%  filter(Catholic < 5)
#   filter(swiss, Catholic < 5)

```

What?? That's it???!!! Why would anyone want to use it?

Well suppose you wanted several operations on the dataframe. Without the pipe you could write

```{r}
swiss_final <- mutate( arrange( filter(swiss, Catholic < 5), Catholic), new_col = Examination + Education)
```

This looks rather chaotic and messy. But with pipes the same expression now becomes

```{r}
swiss_final <- swiss %>% 
  filter(Catholic < 5) %>% 
  arrange(Catholic) %>% 
  mutate(new_col = Examination + Education)
```

This, arguably, looks much more readable. We see that first we apply filter then arrange and then mutate.
